---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
devtools::install()
```

# `afrobarometer`: Load Afrobarometer Data into R

The `afrobarometer` automates laoding Afrobarometer Survey data into the R
environment.  Open access data is downloaded as needed.  The user supplies any
optional restricted access data.  Data is cached in the Afrobarometer data
directory, merged in a master SQLite database in the data directory, and read
into R as a local `tibble`.

- [Data](http://www.afrobarometer.org/data)  
- [Data usage policy](http://www.afrobarometer.org/data/data-use-policy)

## Install

Install `afrobarometer` package

```{r, eval=F}
devtools::install_github("sboysel/afrobarometer")
```

Note that the following packages must be installed.  It may be required that 
some system software packages need to be installed outside of R.

R package   | Dependencies 
------------|-------------
`rgdal`     | `gdal`
`rgeos`     | `geos`
`RMYSqlite` | `sqlite3`

## Required Data

- Open access data: individual responses to survey questionnaires.
- Restricted access data (optional): spatial locations, recently administered
  rounds.
  
The `afrobarometer` package fetches any public access data necessary and merges
the available spatial data with the questionnaire data.  If spatial data is not
provided for a specific round, the final table is simply the questionnaire
data.


## Usage

```{r}
library(afrobarometer)
afrb_dir(path = tempdir())
```

`afrb_dir` sets the local cache directory to `path`, creating subdirectories
and the local database file as needed.  If you have access to spatial location 
data for each round, place the CSV files in the `location` subdirectory of the 
Afrobarometer and name them `Locations_R1.csv`, `Locations_R2.csv`, etc. For 
example, if you have spatial information for rounds 3 and 4, your Afrobarometer
data directory should look like this after running `afrb_dir`:  

```
.
├── locations
│   ├── Locations_R3.csv
│   └── Locations_R4.csv
├── questionnaires
└── afrobarometer.sqlite

2 directories, 3 files
```

Build the database locally

```{r, eval=F}
afrb_build(rounds = c(3, 4), overwrite_db = TRUE)
```

Pull the merged Round 3 data into R 

```{r, eval=F}
r3 <- afrb_round(3)
```

Notes

1. Column names are converted to lowercase on import.

## Citation

This package simply structures the Afrobarometer data for analysis.  Any use of the
Afrobarometer data must comply with it's [Data Usage and Access Policy](http://www.afrobarometer.org/data/data-use-policy).  The data is protected by copywright and the authors request the following citation for any
use of either the open or restricted data:

    Afrobarometer Data, [Country(ies)], [Round(s)], [Year(s)], available at http://www.afrobarometer.org

## TODO

- [ ] Package tests
- [ ] Download codebooks
- [ ] Discuss merging details
