---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
devtools::install()
```

# `afrobarometer`: Load Afrobarometer Data into R

The `afrobarometer` automates laoding Afrobarometer Survey data into the R
environment.  Open access data is downloaded as needed.  The user supplies any
optional restricted access data (i.e. survey respondents georeferenced by 
sampling unit).  Merged data is cached in the Afrobarometer data directory and
read into R as a local `data.frame`.

- [Data](http://www.afrobarometer.org/data)  
- [Data usage policy](http://www.afrobarometer.org/data/data-use-policy)

## Install

```{r, eval=F}
devtools::install_github("sboysel/afrobarometer")
```

## Data

- Open access data: individual responses to survey questionnaires.
- Restricted access data (optional, but must be placed manually): spatial 
  locations, recently administered rounds.
  
The `afrobarometer` package fetches any public access data necessary and merges
the available spatial data with the questionnaire data.  If spatial data is not
provided for a specific round, the final table is simply the questionnaire
data.

## Usage

```{r, eval=FALSE}
library(afrobarometer)
afrb_dir(path = "~/Dropbox/dataset/Afrobarometer")
```

`afrb_dir` sets the local cache directory to `path`, creating subdirectories
and the local database file as needed.  If you have access to spatial location 
data for each round, place the CSV files in the `location` subdirectory of the 
Afrobarometer and name them `Locations_R1.csv`, `Locations_R2.csv`, etc. For 
example, if you have spatial information for rounds 3 and 4, your Afrobarometer
data directory should look like this after running `afrb_dir`:  

```
.
├── locations
│   ├── Locations_R3.csv
│   └── Locations_R4.csv
├── questionnaires
└── codebooks

3 directories, 2 files
```

Build the database locally

```{r, eval=F}
afrb_build(rounds = 3:4, overwrite = TRUE)
```

Open access questionnaire data and codebooks are downloaded as needed and placed in the
Afrobarometer data directory, \code{getOption("afrobarometer.data")}.  If spatial data is
available, it is merged to questionnaire data for each round.  Built datasets are stored in
the subdirectory \code{build} of \code{getOption("afrobarometer.data")}.  The Afrobarometer
directory should now have the following structure:

```
.
├── build
│   ├── afrb_3.rds
│   └── afrb_4.rds
├── locations
│   ├── Locations_R3.csv
│   └── Locations_R4.csv
├── questionnaires
│   ├── merged_r3_data.sav
│   └── merged_r4_data.sav
└── codebooks
    ├── merged_r3_codebook2_0.pdf
    └── merged_r4_codebook3.pdf

4 directories, 8 files
```

Pull the merged Round 3 data into R 

```{r, eval=F}
r3 <- afrb_round(round = 3)
```

## Notes

1. Column names are converted to lowercase on import.
2. As of now, the only variables kepted from the the location files (i.e. Afrobarometer 
  georeferenced data) are the respondent number, latitude, and longitude of the respondent 
  (i.e. respondent's cluster).  These variables are then merged with the full questionnaire
  by respondent number.

## Citation

This package simply structures the Afrobarometer data for analysis.  Any use of the
Afrobarometer data must comply with it's [Data Usage and Access Policy](http://www.afrobarometer.org/data/data-use-policy).  The data is protected by copywright and the authors request the following citation for any
use of either the open or restricted data:

    Afrobarometer Data, [Country(ies)], [Round(s)], [Year(s)], available at http://www.afrobarometer.org

## TODO

- [ ] Package tests (afrobarometer Coverage: 18.28%)
- [x] Download codebooks
- [x] Discuss merging details: take lat + long + respno
- [x] Switch to caching merged data as `rds` in `build` directory.
- [ ] Merge multiple rounds together (Inter-round question correspondence, if possible)
